# Preparation

We have prepared the workshop as much as we could in a language-independent manner. That was not always possible, and where we are going into detail on tools and tricks we have done so catering to R and Python users. If you use a different language, you are very welcome to join our workshop, and we hope you will be able to help us add to the teaching materials with comparable tools you have found.

Please go through this chapter carefully before the workshop, especially the sub-section on _Git_ and the _Final Test!_. It's necessary to have Git working so you can go through the course (relatively) smoothly. 

## Data and/or Code

-   **Bring some code!** We will be working on improving your workflow, so it is nice to have material to work with. Consider that we will need something that is small enough that you can edit/play with it during the workshop, but also comprehensive so that someone else may be able to reproduce your project on their computer (i.e.: it is nice if there is an output for someone to see at the end!). *NB: If you will not have code to work on, you can still take advantage of the workshop, but please contact your instructor before the course.*

-   **Optional: if your data cannot be public, fake it!** We will be exchanging repositories at the end of the day, and try to reproduce each other's work. This can be problematic if your data cannot be shared. If the structure of the data can be put online, you could simulate the dataset (including covariates, so simple models have similar outcomes) using a package like [fakeR](https://cran.r-project.org/web/packages/fakeR/index.html), which simulates fake data within the same structure. Or check out or [this post on conjurer](https://www.r-bloggers.com/generate-synthetic-data-using-r/).

## Terminal

Some of the work we will do will take place over the command line, for which you will need a terminal.

-   Mac OS and Linux have easily accessible terminals (the program is usually called _Terminal_, so that's easy).

-   Windows' default terminal (Command Prompt) uses different commands, so we do not recommend it for this workshop. Instead, as you install Git (next sub-section), you can get the Git Bash terminal with your installation.

If you have no experience with the terminal at all, we recommend practicing with it for some basic folder navigation to start. [This 8.5 minute video](https://www.youtube.com/watch?v=j6vKLJxAKfw) is a good introduction.

## Git

### Coding Platforms

Sign up for a [GitHub](https://github.com/join?source=header-home), [GitLab](https://gitlab.com/users/sign_up), or [BitBucket](https://bitbucket.org/account/signup/) account. 

Please note our workshop materials and demonstrations will be in GitHub, but most things we will show you have easy parallels on the other networks, so you can likely just follow along.

### Install Git

You first want to double-check if Git is already installed on your computer.

Go to your terminal and enter `which git` to request the path to your Git executable:

```{r eval=FALSE}
which git
## /usr/local/bin/git
```

and then `git --version` to see its version:

```{r eval=FALSE}
git --version
## git version 2.31.1
```

If the terminal returns a path and version, that means you have Git and you can skip this installation step! That being said, you might want to ensure your Git is the latest version.

If you don't see a path and version, but something like `git: command not found` - then you definitely need to continue with installation. Follow the the instructions relevant to your operating system.

* **Windows:** Go to [gitforwindows.org](https://gitforwindows.org). Make sure you include **Git Bash** in your installation!

* **Mac:** Go to [git-scm.com](http://git-scm.com/downloads) and install the the latest version available. 

* **Linux:** Go the Linux-specific page of [git-scm.com](https://git-scm.com/download/linux). There are commands specific to various package managers. 

### `git config`

Now that installation is done, it's time to introduce yourself to Git!

Go to your terminal and type the following, one after the other:

```{r eval=FALSE}
git config --global user.name 'Your Name'
git config --global user.email 'yourname@example.com'
git config --global --list
```

**Important:** make sure you use the email associated with your **GitHub** account! 

The first two commands won't return any information, the last one will confirm if Git received your introduction alright. That is, your name and email will be returned.

### Set up SSH keys 

SSH keys provide a more secure way of logging onto servers than relying on passwords alone. It's similar to how two-factor authentication provides an extra layer of security over regular passwords/passcodes.

There are number of steps involved in this setup, but we're going to summarize it here - so you don't get lost:

1. Check for existing SSH keys.
2. Generate new SSH keys.
3. Add private key to the ssh-agent.
4. Add public key to your GitHub/GitLab/BitBucket account.
5. Test your SSH connection.

These instructions are borrowed heavily from [Happy Git with R](https://happygitwithr.com/), GitHub's user guides, and [NLeSC Small Scale Initiative - Introduction to GitVersion Control with Git](https://bvreede.github.io/presentations/presentations/2021-05-18_Git-intro-escience.html#1).

#### Check for existing SSH keys.

Go to your terminal and type the following command to list existing SSH keys:

```{r eval=FALSE}
ls -al ~/.ssh
```

If you have key pairs already, they may look like the following:

* `id_rsa.pub` (public key) & `id_rsa` (private key)

* `id_ecdsa.pub` (public key) & `id_ecdsa` (private key)

* `id_ed25519.pub` (public key) & `id_ed25519` (private key)

If you're happy to stick with your existing keys, then you can jump to step 3 which is about adding the private key to the ssh-agent.

#### Generate new SSH keys

Go to your terminal and type the following command, substituting "USEFUL-LABEL" with something like your GitHub username and/or your device.  

```{r eval=FALSE}
ssh-keygen -t ed25519 -C "USEFUL-LABEL"
```

Accept the proposal to save the key in the default file location. Just press Enter here:

```{r eval=FALSE}
> Enter a file in which to save the key (/Users/you/.ssh/id_ed25519): [Press enter]
```

If you're new to setting up SSH keys, then skip the option to enter a passphrase. It's a best practice, but you can come back to it after the workshop. Just leave the passphrase option empty and press Enter:

```{r eval=FALSE}
> Enter passphrase (empty for no passphrase):
```

Here's an example of how it should have played out, if everything went smoothly:

```{r eval=FALSE}
jenny@2015-mbp ~ $ ssh-keygen -t rsa -b 4096 -C "USEFUL-COMMENT"
Generating public/private rsa key pair.
Enter file in which to save the key (/Users/jenny/.ssh/id_rsa):     
Enter passphrase (empty for no passphrase): 
Enter same passphrase again: 
Your identification has been saved in /Users/jenny/.ssh/id_rsa.
Your public key has been saved in /Users/jenny/.ssh/id_rsa.pub.
The key fingerprint is:
SHA256:ki0TNHm8qIvpH7/c0qQmdv2xxhYHCwlpn3+rVhKVeDo USEFUL-COMMENT
The key's randomart image is:
+---[RSA 4096]----+
|      o+   . .   |
|     .=.o . +    |
|     ..= + +     |
|      .+* E      |
|     .= So =     |
|    .  +. = +    |
|   o.. = ..* .   |
|  o ++=.o =o.    |
| ..o.++o.=+.     |
+----[SHA256]-----+
```

#### Add private key to the ssh-agent

While the actions are similar, the specific commands for this step vary between operating systems. Therefore, we'll link you to the relevant GitHub guide for this step: [Generating a new SSH key and adding it to the ssh-agent](https://docs.github.com/en/github/authenticating-to-github/connecting-to-github-with-ssh/generating-a-new-ssh-key-and-adding-it-to-the-ssh-agent#adding-your-ssh-key-to-the-ssh-agent)

Make sure you're following the instructions for the right operating system and then jump to the sub-section on _Adding your SSH key to the ssh-agent_.

Effectively, what this step does is to 'activate' your private key.

#### Add public key to your GitHub/GitLab/BitBucket account

* Open your public SSH key (the file ending with .pub) in a text editor and copy the whole key. Make sure to avoid copying any newlines or whitespace. 

* On GitHub, click on your profile picture and go to _Settings_ -> _SSH & GPG keys_. 

* Click on _New SSH key_.

* Paste your public key in the box and give it an informative label, maybe the same as the "USEFUL-LABEL" you used when generating the keys.

#### Test your SSH connection

Open your terminal and type in the following command:

```{r eval=FALSE}
ssh -T git@github.com
```

You may see a warning which includes an RSA key fingerprint. Make sure this fingerprint matches that of GitHub's, you can double-check that here: [GitHub's SSH key fingerprints](https://docs.github.com/en/github/authenticating-to-github/keeping-your-account-and-data-secure/githubs-ssh-key-fingerprints)

If the fingerprint matches, type _yes_ if you were asked if wanted to continue connecting.

If all goes well, you'll see something like the following:

```{r eval=FALSE}
> Hi username! You've successfully authenticated, but GitHub does not provide shell access.
```

Make sure it's your username that terminal returns!

#### Troubleshooting 

GitHub has a whole page on [Troubleshooting SSH](https://docs.github.com/en/github/authenticating-to-github/troubleshooting-ssh). If you run into issues, you should double-check if your errors are covered here. 

We have collected some commonly seen errors:

* **You have made the key pair and copied the public key to GitHub, but SSH will not connect.**
  Are you sure that you [turned on the agent](https://docs.github.com/en/github/authenticating-to-github/generating-a-new-ssh-key-and-adding-it-to-the-ssh-agent#adding-your-ssh-key-to-the-ssh-agent)?
  
* You recieve an error message like ["Error: Permission denied (publickey)"](https://docs.github.com/en/github/authenticating-to-github/troubleshooting-ssh/error-permission-denied-publickey). Follow the link to GitHub guide on this issue.

* Did you add the public key to GitHub?

* Toggle between SSH & HTTPS 

* Did you configure Mac OS Sierra or High Sierra to persistently store your passphrase in the keychain?

* **You can run commands like `ssh-keygen` or `eval`, but they do not return any information, not even an error message.**
  If you are on Windows and have have Citrix workspace installed, there is a known bug where Citrix interferes with the SSH key generation process.
  - Uninstall Citrix workspace
  - Restart your computer
  - Restart the SSH key generation tutorial.
  - If, after setting up SSH key pairs, you want to reinstall Citrix, be sure to do so via the Microsoft store.

### If SSH fails, HTTPS!

If you find yourself despairing over the SSH key setup, don't worry! HTTPS is a sufficient back-up!

You will need to create a [Personal Access Token from Github](https://docs.github.com/en/github/authenticating-to-github/creating-a-personal-access-token) to identify yourself over the HTTPS protocol:

* Click on your profile icon (top right side of your screen)

* Go to _Settings_ -> _Developer Settings_ -> _Personal Access Tokens_

* Create a new token, and select the ‘repo’ scope for its access.

* Copy the token and use it when Git asks for a password when connecting to Github.

See the [Github website for more info](https://docs.github.com/en/github/authenticating-to-github/creating-a-personal-access-token).

#### Troubleshooting Personal Access Tokens

You can only view a Personal Access Token once, and they are impossible to type. However, you can store it in your system so it can be used to authenticate you without your having to re-enter it.

* If you are on Windows, go to the terminal and type `git config --global credential.helper wincred`

* If you are on a different system, and git does not cache your credentials automatically, [here is a Github tutorial to fix that](https://docs.github.com/en/github/getting-started-with-github/caching-your-github-credentials-in-git).

## Final Test!

If you've set up your SSH keys alright, we can try going ahead and seeing whether you can _push_ and _pull_ from GitHub successfully. If you're going the HTTPS route due to issues, no worries! Just make sure your switch out the SSH clone URL for the HTTPS clone URL. 

This is something we'll do during the workshop as well, but trying it out beforehand reduces chances of errors while the course is ongoing. 

The sequence of steps is as follows:

1. Create a test repository on GitHub.

2. Clone the test repository. 

3. Make local changes.

4. _Push_ the local changes to GitHub. This is where the magic happens (or not)!

5. If all went well, you can refresh the webpage with your GitHub repository and you'll see your local changes.

6. Make changes online.

7. _Pull_ the online changes to your computer, so your local repository is up to date.

8. If all went well, then your online changes are available locally as well. It also means you're fully set up with Git & GitHub!

#### Create a test repository on GitHub

* Log into GitHub.

* Click the green _New_ (repository) button.

* You can name your repository as `testrepo`.

* Make sure it's a Public repository.

* Click _YES_, when asked if you want to initialize the repository with a README.

* Click the green _Create repository_ button.

* Copy the SSH URL for cloning via the green _Clone or Download_ button.

#### Clone the test repository 

Go to your terminal and type the following command, substituting the SSH clone URL you copied in the previous step after `git clone`:

```{r eval=FALSE}
git clone git@github.com:username/testrepo.git
```

This should create a local copy of your repository as a folder with files contained within.

#### Make local changes

Navigate to the folder of your repository and open the README file. Make an edit, such as _This is a line I'm adding offline to my local copy_. Save the file and close it.

#### Push the local changes to GitHub

Go to your terminal and type the following command(s):

```{r eval=FALSE}
git add .
```

This adds all changed files to a 'staging area'.

```{r eval=FALSE}
git commit -m "insert-understandable-message-here"
```

Now you're officially 'committing' the changes you made. You want to add a understandable commit message so you have a clear record of your changes.

```{r eval=FALSE}
git push
```

With this command, you push the local changes and commit message to your repository online. 

This is where errors can pop up if the SSH set-up didn't go as smoothly as expected. Troubleshoot the error messages the best you can!

If everything went smoothly, you can refresh the webpage with GitHub repository and you should see that your changes have been synced.

#### Make changes online

Now we want to go the other way around, open your README file on GitHub and make an edit online. You can do this by: 

* Clicking on the README file and the pencil icon to open edit mode. 

* Add a sentence to your README like _This is a line I'm adding from GitHub.com_. 

* When saving, you'll have to provide a meaningful commit message again.

#### Pull from GitHub

Now go to your terminal and type the following command:

```{r eval=FALSE}
git pull
```

If all goes well, you'll see some activity in the terminal which suggests the changes are being downloaded. You can then check out your README file and see whether your online change has been synced.

#### Wrap-Up 

Did it work? You did it! You pushed and pulled from GitHub successfully, woohoo!
