# README Files

### Video

```{r echo=FALSE}
vembedr::embed_url("https://vimeo.com/464027978")
```

### Slides

```{r echo=FALSE}
xaringanExtra::embed_xaringan(url = "slides/slides_documentation.html#16", ratio = "4:3")

```
[Slides](slides/slides_documentation.html#16)

### Exercise

Edit your README file.
Take your time. This is important!

Make sure to include the following information:

* What does your project do?

* How does the user access your project? (E.g. download, or clone with `git clone`...)

* How does the user call the main script(s) that should be executed?

* And perhaps any other elements you are inspired to add (check out the examples!)

Do you want to truly understand the importance of decent installation information?
[Read this experience by Elisabeth Bik](https://twitter.com/MicrobiomDigest/status/1283082285097422848), trying to install image forensics software.
Look at the [README of the software](https://github.com/GuidoBartoli/sherloq), and see what you can learn.

_NB: We will have another chance to take a look at dependencies and prerequisites, but if you have opportunity to add these here: please do!_
