# Robustness

### Video

```{r echo=FALSE}
vembedr::embed_url("https://vimeo.com/463636970")
```

### Slides

```{r echo=FALSE}
xaringanExtra::embed_xaringan(url = "slides/slides_code-quality.html#21", ratio = "4:3")
```
[Slides](slides/slides_code-quality.html#21)

### Exercise

* Identify assumptions in your code
  
  * What assumptions/expectations exist on your data or (user) input?
  
  * What assumptions/expectations exist on the input of (a) function(s)?

Based on the assumptions/expectations you were able to identify, work on increasing the robustness of your code.
Choose one:

* Make the input/data assumptions explicit
  
  * Option 1: Explicitly state assumptions on data or input in your README.md.
  
  * Option 2: Write a piece of code that tests the validity of data/input, and reports an error if the expectations are not met.

* Test the input for a function. Modify the code inside your function to:
  
  * check the value of the arguments passed to your function using if/else statements;
  
  * raise an error in case an argument is out of the range of acceptable values.
