# Version control

### Video

```{r echo=FALSE}
vembedr::embed_url("https://vimeo.com/463264170")
```

### Slides

```{r echo=FALSE}
xaringanExtra::embed_xaringan(url = "slides/slides_project-setup.html", ratio = "4:3")
```
[Slides](slides/slides_project-setup.html)

### Exercise

* Follow the steps in the video (or [on these slides](slides/slides_project-setup.html#17)) to turn your folder into a git repository.

* Make a remote version of your project on GitHub!

* Please note: are there (temporary) files you do not wish to track?
  Add them to the `.gitignore` file.
  Consider a `.gitignore` template for your language: examples on [this github repo](https://github.com/github/gitignore).

* Can you use Git and push to Github from your IDE?

* Experiment with editing and committing on GitHub itself. You can then 'download' your code to your local repository using `git pull`.

* Optional: What happens if you edit the same file online and locally, and try to push/pull?
  (Hint: this often causes a 'merge conflict', which is no fun to experience.
  Going through it today means we can assist you if necessary!)